---
title: "Machine Learning class short presentation - Kmeans"
output:
  xaringan::moon_reader:
    seal: false
    self_contained: true  ## otherwise you need the entire folder to see the slides correctly
    yolo: false
    css: ["css_template/middlebury.css", "css_template/tamu-fonts.css"] ## Note that if you don't specify the .css after the file name here, xaringan will import and renew the template at each run in ./libs/remark-css/
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: [top, left, inverse]
---

```{r, echo=F}
knitr::opts_chunk$set(message = F, 
                      warning = F)
```



class: inverse, center, middle


# Question 1

Load the Titanic data set and run the K-Means algorithm to cluster the passengers into 2 clusters (K = 2). Do not use the attribute Survived.


How much do these two clusters and the Survived and NotSurvived subgroups overlap?
 

Experiment with different subsets of the given features. You can add some new feature(s).

---

# Set up the project

</br>
</br>

* Libraries

```{r}
rm(list = ls())
library(tidyverse)
library(ggpubr) # to plot the kmean
library(factoextra) # to get the coordinates of the PCA
```

</br>
</br>

* Load the dataset

```{r,}
titanic <- read_csv("data/train.csv")
```

---

.center[**Load the Titanic data set and run the K-Means algorithm to cluster the passengers into 2 clusters (K = 2). Do not use the attribute Survived.**]

</br>
</br>

* K-mean algorithm works only on numeric data. Thus, we have to transform binary character variables into numeric ones (*e.g* sex) and take off the qualitative variables.

```{r}
## Prepare the data
clust_data <- 
  titanic %>% 
  mutate(Sex = ifelse(Sex == "male", 1, 0)) %>%   ## transforme sex into numeric binary variable
  select_if(Negate(is.character)) %>%             ## filter the columns that are numeric
  na.omit()
```

* Perform the clustering on `r colnames(clust_data %>% select(-Survived, -PassengerId))`

```{r}
set.seed(123)

kmean1 <- 
  kmeans(clust_data %>% select(-Survived, -PassengerId), centers = 2, nstart = 1000)

```

---

.center[**How much do these two clusters and the Survived and NotSurvived subgroups overlap?**]

* In order to represent the clusters, I perform a PCA to reduce the feature space

```{r}
set.seed(123)

## Reduce the space with PCA
pca <- prcomp(clust_data %>% select(-Survived, -PassengerId), scale = T)

## Extract coordiantes of the passengers:
indcoord <- as.data.frame(get_pca_ind(pca)$coord)

## Add the cluster to the coord:
indcoord$cluster <- factor(kmean1$cluster)

## Add the survived info
indcoord$surv <- clust_data$Survived

# Percentage of variance explained by dimensions
eigenvalue <- round(get_eigenvalue(pca), 1)
```


---


```{r, fig.width=10, fig.height=6, fig.align="center"}
## PLot it
indcoord %>% 
  ggplot(aes(Dim.1, Dim.2, color = cluster)) +
  geom_point(aes(shape = as.factor(surv)), size = 3)+
  stat_chull(aes(fill = cluster), geom = "polygon", alpha = .5)+
  xlab(paste0("Dim 1 (", eigenvalue$variance.percent[1], " %)"))+
  ylab(paste0("Dim 2 (", eigenvalue$variance.percent[2], " %)"))+
  ggtitle("Kmean representation")+
  theme_bw()
```



---



.center[**How much do these two clusters and the Survived and NotSurvived subgroups overlap?**]

</br>
</br>

* Merge the clustering output to the dataset

```{r}
clust_data <- cbind(clust_data, cluster = kmean1$cluster)
```

</br>
</br>

* Look how `Survived` and `cluster` interfere:

```{r}
table(clust_data[c("Survived", "cluster")])
```

.center[That's pretty bad..]

---


.center[**Experiment with different subsets of the given features. You can add some new feature(s).**]

* With `Age, Sex, Pclass`

```{r}
kmean2 <- kmeans(clust_data %>% select(Age, Sex, Pclass), centers = 2, nstart = 1000)
clust_data <- cbind(clust_data, cluster2 = kmean2$cluster)
table(clust_data[c("Survived", "cluster2")])
```
* With `Age, Sex, Pclass, Fare`

```{r}
kmean3 <- kmeans(clust_data %>% select(Age, Sex, Pclass, Fare), centers = 2, nstart = 1000)
clust_data <- cbind(clust_data, cluster3 = kmean3$cluster)
table(clust_data[c("Survived", "cluster3")])
```

---

# Conclusion

</br>
</br>

* K-mean doesn't seem really efficient in predicting the survival of the passengers

</br>
</br>

* In my opinion, a logistic regression would be more efficient

